---
title: "Data Science for Business: Homework Assignment"
author: "Gyurmalin"
date: "February 21, 2016"
output: html_document
fontsize: 14p
---

******



### **Bank Marketing Data Set**


#### Downloaded from UCI Machine Learning Repository

******
#### Source: http://archive.ics.uci.edu/ml/datasets/Bank+Marketing.

#### This dataset is public available for research. The details are described in S. Moro, P. Cortez and P. Rita. A Data-Driven Approach to Predict the Success of Bank Telemarketing. Decision Support Systems, Elsevier, 62:22-31, June 2014

******

#### Introduction

This dataset provides unformation of telemarketing campaigns pursued by a Portuguese bank. Aim of current analysis is to predict whether a client would subscribe for a term deposit (yes/no). There are 20 input variables and a binary output variable.

**For this assignment I decided to mark all code chunks in the documentation in order to track processes step-by-step.**

******

#### Set libraries


```{r
opts_chunk$set(warning=FALSE, error=FALSE, 
message=FALSE)}

library(readr)
library(dplyr)
library(reshape2)
library(ggplot2)

library(knitr)
opts_knit$set(root.dir = "../")
```

******

#### Read data and check its structure

As mentioned above, we have 20 predictors and 1 output (altogehther 21 features).
There are altogether 41188 observations.
```{r}
bank <- read.csv("C:/Users/Szaffi/Desktop/Gyurma/Data_Science_for_Business/bank-additional/bank-additional-full.csv", sep=";")
```

Our variables are the following:
```{r}
colnames(bank)
```

Let's have a look at the structure of our data. As mentioned above, we have 20 predictors and 1 output (altogehther 21 features).
There are altogether 41188 observations.
```{r}
dim(bank)
str(bank)
```

And check some base descriptive statistics on raw data set:
```{r}
summary(bank)
```

#### Have a quick look at data itself

```{r}
head(bank)
```
